\bf Range Analysis for the IonMonkey JavaScript engine

### 2012-05-01 project-writings/paper.tex
our pseudocode is valid, syntax highlighted python

\documentclass{article}
\usepackage{fullpage}
%\usepackage{times}
\usepackage{url}
\usepackage{amsmath,amsthm,amssymb,color}
%\usepackage{multicol}
\usepackage{float}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
%\usepackage[all]{hypcap}

\newcommand{\Squash}{\mathsf{Squash}}
\newcommand{\Clamp}{\mathsf{Clamp}}
\newcommand{\Range}{\mathsf{Range}}
\let\x\cdot

@@

Thus we developed a quite simple and in practice efficient algorithm to
add $\beta$ nodes to an SSA source program by taking
advantage of data structures common to modern SSA based compilers
\footnote{The method for removing $\beta$ nodes is so trivial that we
  could not claim with a straight face that we ``developed an
  algorithm'' to perform it.} .  Pseudocode for the algorithm is shown
in Figure \ref{fig:add_betas}. The algorithm traverses the control
flow group, looking for blocks that have been entered by a jump
conditioned on a relational operation. When we find a variable being
compared to a constant, we compute a range and use it to create a new
$\beta$ node. We want to propagate this range information to all uses
of the variable where the branch is guaranteed to have been taken;
this is accomplished by replacing all uses of original instruction
that are dominated by the new $\beta$ instruction with the $\beta$. To
do this, we take advantage of IonMonkey's use chains to efficiently
iterate over all uses of an instruction.

@@

\begin{figure}[ht]
\begin{center}
\begin{verbatim}
function addBetaNodesToFunc(cfg):
  for each b in cfg.basic-blocks():
    if b.numPredecessors() == 1 and isBranch(b.pred(0).lastInstruction())
      test = b.pred(0).lastInstruction()
      # find whether we are the true or false branch
      direction = (b == test.trueTarget())
      if test is of the form <x RELOP c>:
          r = computeRange(RELOP, c, direction)
          x' = b.insertFront(<Beta(x, r)>)
          replaceDominatedUses(x, x')

# Replace all uses of the instruction x that are dominated by
# the instruction x' with x'
function replaceDominatedUses(x, x'):
  for each use in x.uses():
    ins = use.instruction()
    if ins != x' and isDominatedUse(x', use):
      replaceUseWith(use, x')

function isDominatedUse(x', use):
  ins = use.instruction()
  if ins.isPhiNode():
    # Get the predecessor block that corresponds to the
    # argument of the phi node we are looking at
    phi-arg-pred-block = ins.block().pred(use.index())
    return blockDominates(x'.block(), phi-arg-pred-block)
  else:
    return blockDominates(x'.block(), x.block())

function blockDominates(b1, b2):
  high = b1.domIndex() + b1.numDominated()
  low  = b1.domIndex()
  return b2.domIndex() >= low && b2.domIndex <= high
\end{verbatim}
\end{center}
\caption{Algorithm for inserting beta nodes}
\label{fig:add_betas}
\end{figure}

>>>>

\documentclass{article}
\usepackage{fullpage}
%\usepackage{times}
\usepackage{url}
\usepackage{amsmath,amsthm,amssymb,color}
%\usepackage{multicol}
\usepackage{float}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{courier}
\usepackage{listings}
%\usepackage[all]{hypcap}

\newcommand{\Squash}{\mathsf{Squash}}
\newcommand{\Clamp}{\mathsf{Clamp}}
\newcommand{\Range}{\mathsf{Range}}

\let\x\cdot
\lstset{
  language=Python,
  showstringspaces=false,
  tabsize=4,
  basicstyle=\ttfamily,
  keywordstyle=\bfseries\ttfamily,
  captionpos=b
}

@@

Thus we developed a quite simple and in practice efficient algorithm to
add $\beta$ nodes to an SSA source program by taking
advantage of data structures common to modern SSA based compilers
\footnote{The method for removing $\beta$ nodes is so trivial that we
  could not claim with a straight face that we ``developed an
  algorithm'' to perform it.} .  Pseudocode for the algorithm is shown
in Listing \ref{lst:add_betas}. The algorithm traverses the control
flow group, looking for blocks that have been entered by a jump
conditioned on a relational operation. When we find a variable being
compared to a constant, we compute a range and use it to create a new
$\beta$ node. We want to propagate this range information to all uses
of the variable where the branch is guaranteed to have been taken;
this is accomplished by replacing all uses of original instruction
that are dominated by the new $\beta$ instruction with the $\beta$. To
do this, we take advantage of IonMonkey's use chains to efficiently
iterate over all uses of an instruction.

@@

\begin{lstlisting}[language=Python,
                   caption={Pseudocode algorithm for inserting beta nodes},
                   label={lst:add_betas}]
def addBetaNodesToFunc(cfg):
  for b in cfg.basic_blocks():
    if b.numPredecessors() == 1 and isBranch(b.pred(0).lastInstruction()):
      test = b.pred(0).lastInstruction()
      # find whether we are the true or false branch
      direction = (b == test.trueTarget())
      if test.op() in RELOPs:
          r = computeRange(test.op(), c, direction)
          x_prime = b.insertFront(Beta(x, r))
          replaceDominatedUses(x, x_prime)

# Replace all uses of the instruction x that are dominated by
# the instruction x_prime with x_prime
def replaceDominatedUses(x, x_prime):
  for each use in x.uses():
    ins = use.instruction()
    if ins != x_prime and isDominatedUse(x_prime, use):
      replaceUseWith(use, x_prime)

def isDominatedUse(x_prime, use):
  ins = use.instruction()
  if ins.isPhiNode():
    # Get the predecessor block that corresponds to the
    # argument of the phi node we are looking at
    phi_arg_pred_block = ins.block().pred(use.index())
    return blockDominates(x_prime.block(), phi_arg_pred_block)
  else:
    return blockDominates(x_prime.block(), x.block())

def blockDominates(b1, b2):
  high = b1.domIndex() + b1.numDominated()
  low  = b1.domIndex()
  return b2.domIndex() >= low and b2.domIndex <= high
\end{lstlisting}
